---
title: "homework 4"
author: "Jeffrey Zhuohui Liang"
date: "3/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(pdp)
library(lasso2)
library(klaR)
library(earth)
library(rpart)
library(party)
library(doParallel)
library(caret)
library(DALEX)
library(tidyverse)

knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = T,
  warning = F,
  cache = T
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  digits = 3
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(123123)
knitr::opts_chunk$set(echo = TRUE)
```

# 1
```{r data_step}
data(Prostate)

Prostate = Prostate %>% 
  janitor::clean_names()

trainindex = createDataPartition(Prostate$lpsa,p=0.8,list = F)

X_tr = model.matrix(lpsa~.,Prostate[trainindex,])[,-1]

Y_tr = Prostate[trainindex,"lpsa"]


X_ts = model.matrix(lpsa~.,Prostate[-trainindex,])[,-1]

Y_ts = Prostate[-trainindex,"lpsa"]

ctrl = trainControl(method = "repeatedcv",number = 5, repeats = 5)

pre = c("center","scale")

```


```{r single_tree}
sng_tree = rpart(lpsa ~ .,
                     data = Prostate,
                     subset = trainindex,
                     control = rpart.control(cp = 0))

plotcp(sng_tree)

cpTable = sng_tree$cptable

minErr = which.min(cpTable[,4])

sng_tree_1se = prune(sng_tree,
                     cp=cpTable[cpTable[,4]<cpTable[minErr,4]+cpTable[minErr,5],1][1])

sng_tree_min = prune(sng_tree,cp = cpTable[minErr,1][1])

last(sng_tree_min$cptable[,2])

last(sng_tree_1se$cptable[,2])

```

The number of node of minumum RMSE tree is the same as  1se tree's number of node.

# 2

```{r}
plot(partykit::as.party(sng_tree_1se))
```

For those observation whom lcavol<-0.479 will go into terminal node 3, which has a mean response near 0.5 with 8 observations.

# 3
```{r bagging}
cl = parallel::makePSOCKcluster(5)
doParallel::registerDoParallel(cl)

sng_tree = train(X_tr,
                 Y_tr,
                 method = "rpart",
                 tuneGrid = expand.grid(cp = exp(seq(-3,-5,len = 20))),
                 preProcess = pre,
                 trControl = ctrl
                 )

stopCluster(cl)

bagging = 
  train(X_tr,
        Y_tr,
        method = "ranger",
        tuneGrid = expand.grid(mtry = 8,
                               splitrule = "variance",
                               min.node.size = 1:20),
        metric = "RMSE",
        trControl = ctrl,
        preProcess = c("center","scale"))

bag_explain = DALEX::explain(bagging,
                             label = "bagging",
                             data=X_tr,
                             y= Y_tr,
                             verbose = F)

bag_imp = DALEX::model_parts(bag_explain)

plot(bag_imp)

parallel::stopCluster(cl)

ggplot(bagging,highlight = T)
```

The order of variable importance are `r bag_explain$data %>% as_tibble() %>% names()`.

# 3
```{r rf}
cl = parallel::makePSOCKcluster(5)
doParallel::registerDoParallel(cl)

random_forest = 
  train(X_tr,
        Y_tr,
        method = "ranger",
        tuneGrid = expand.grid(mtry = 1:7,
                               splitrule = "variance",
                               min.node.size = 1:20),
        metric = "RMSE",
        trControl = ctrl,
        preProcess = c("center","scale"))

rf_explain = DALEX::explain(random_forest,
                             label = "random forest",
                             data=X_tr,
                             y= Y_tr,
                            verbose = F)

rf_imp = DALEX::model_parts(rf_explain)

plot(rf_imp)

parallel::stopCluster(cl)
```

The order of variable importance are `r rf_explain$data %>% as_tibble() %>% names()`.

# 4
```{r ada}
t1 = Sys.time()
cl = parallel::makePSOCKcluster(5)
doParallel::registerDoParallel(cl)

adab =   train(X_tr,
        Y_tr,
        method = "gbm",
        tuneGrid = expand.grid(n.trees = c(1e+3,2e+3,3e+3),
                               interaction.depth = 1:5,
                               shrinkage = seq(0.01,0.05,len=3),
                               n.minobsinnode = 1:10),
        metric = "RMSE",
        trControl = ctrl,
        preProcess = c("center","scale"))

adab_explain = DALEX::explain(adab,
                             label = "random forest",
                             data=X_tr,
                             y= Y_tr,
                            verbose = F)

adab_imp = DALEX::model_parts(adab_explain)

plot(adab_imp)

parallel::stopCluster(cl)

runt = system.time() - t1
```

The order of variable importance are `r adab_explain$data %>% as_tibble() %>% names()`.

```{r}
reg_rsmp = 
  resamples(
    list(
      tree = sng_tree_min,
      bagging = bagging,
      random_forest = random_forest,
      boost  = adab
    )
  )
```

